---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
# The base pipe needs R >= 4.1; older versions run the magrittr chunks.
new_r <- getRversion() >= "4.1"
```

```{r setup}
library(shinyfilters)
```

```{r, eval = !new_r, include = !new_r}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`as_filters()` turns a data frame into a set of filters. Printing it shows the
input each column gets, based on its type:

```{r, eval = new_r, include = new_r}
nyc_flights |>
	as_filters()
```

```{r, eval = !new_r, include = !new_r}
nyc_flights %>%
	as_filters()
```

Arguments to `as_filters()` apply to every column. For example, `slider = TRUE`
uses sliders for numeric columns:

```{r, eval = new_r, include = new_r}
nyc_flights |>
	as_filters(slider = TRUE)
```

```{r, eval = !new_r, include = !new_r}
nyc_flights %>%
	as_filters(slider = TRUE)
```

## Choose inputs per column

`with_filter()` sets the input for individual columns. Name the columns, or
select them with tidyselect:

```{r, eval = new_r, include = new_r}
filters <- nyc_flights |>
	as_filters() |>
	with_filter(origin = "radio", carrier = "selectize") |>
	with_filter(where(is.numeric), "slider")

filters
```

```{r, eval = !new_r, include = !new_r}
filters <- nyc_flights %>%
	as_filters() %>%
	with_filter(origin = "radio", carrier = "selectize") %>%
	with_filter(where(is.numeric), "slider")

filters
```

Inputs are keywords like `"radio"` or shiny functions like
`shiny::radioButtons`. See `?with_filter` for the full list.

## Use in a shiny app

`filterInput()` creates the inputs:

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
	sidebarLayout(
		sidebarPanel(filterInput(filters)),
		mainPanel()
	)
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```
