---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
library(shinyfilters)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`shinyfilters()` turns a data frame into a set of filters. Printing it shows the
input each column gets, based on its type:

```{r}
shinyfilters(nyc_flights)
```

## Set defaults for every column

Arguments to `shinyfilters()` apply to applicable columns. `range = TRUE` uses 
a date range for date columns, `selectize = TRUE` uses selectize inputs instead
of select inputs, and `multiple = TRUE` is an argument passed to
`selectizeInput`.

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE, multiple = TRUE)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE, multiple = TRUE)

filters
```

## Choose inputs by keyword

Use `with_filters()` to set the input for individual columns. The 
quickest way is by using a keyword like `"slider"`. See `?with_filter` for the 
full list of keywords.

To choose the columns, you have three options:

1. Name the columns (`distance = "slider"`)
2. Identify them with tidyselect (`where(is.numeric)`), 
3. Use `across_filters()`

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_filter(where(is.numeric), "slider")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_filter(where(is.numeric), "slider")

filters
```

### dplyr

Some `dplyr` verbs are also supported:

* `dplyr::across()`
* `dplyr::mutate()`
* `dplyr::pull()`
* `dplyr::select()`
* `dplyr::transmute()`

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = requireNamespace("dplyr", quietly = TRUE), include = requireNamespace("dplyr", quietly = TRUE), message = FALSE}
library(dplyr)
```

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(across(where(is.numeric), "slider"))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(across(where(is.numeric), "slider"))
```

## Choose inputs by shiny function {.tabset}

A shiny input function works in place of its keyword: `radioButtons`
gives the same input as `"radio"`:

### shinyfilters {#shiny-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_filter(origin = radioButtons)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_filter(origin = radioButtons)

filters
```

### dplyr {#shiny-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(origin = radioButtons)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(origin = radioButtons)
```

## Choose inputs by custom function {.tabset}

Any other input function works too, such as `shinyWidgets::pickerInput`. It
receives the arguments shinyfilters computes for the column's type, here the
`choices` of a character column, so it must accept them.

When you supply a custom function, use `as_filter()` so that shinyfilters knows
which update function to use.

### shinyfilters {#custom-shinyfilters}

`r if (!requireNamespace("shinyWidgets", quietly = TRUE)) "Install shinyWidgets to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE)}
filters <- filters |>
	with_filter(c(carrier, dest), as_filter(
		shinyWidgets::pickerInput,
		.update_fn = shinyWidgets::updatePickerInput
	))

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE)}
filters <- filters %>%
	with_filter(c(carrier, dest), as_filter(
		shinyWidgets::pickerInput,
		.update_fn = shinyWidgets::updatePickerInput
	))

filters
```

### dplyr {#custom-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE) || !requireNamespace("shinyWidgets", quietly = TRUE)) "Install dplyr and shinyWidgets to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE)}
filters <- filters |>
	mutate(across(c(carrier, dest), as_filter(
		shinyWidgets::pickerInput,
		.update_fn = shinyWidgets::updatePickerInput
	)))

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE)}
filters <- filters %>%
	mutate(across(c(carrier, dest), as_filter(
		shinyWidgets::pickerInput,
		.update_fn = shinyWidgets::updatePickerInput
	)))

filters
```

## Set an input's arguments {.tabset}

Sometimes you want to modify the arguments that are sent to the input. Use
`as_filter()` for this, inside `with_filter()`.

* Use `.x` to refer to the column being passed to the input.
* `na.rm = TRUE` will be added to the call, if possible.

### shinyfilters {#args-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_filter(
		dep_delay = as_filter(sliderInput, value = range(.x))
	)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_filter(
		dep_delay = as_filter(sliderInput, value = range(.x))
	)

filters
```

### dplyr {#args-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(
		dep_delay = as_filter(sliderInput, value = range(.x))
	)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(
		dep_delay = as_filter(sliderInput, value = range(.x))
	)
```

## Add a namespace

`with_ns()` prefixes every input id with a namespace, so the filters can be
used in a shiny module:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_ns("sidebar-mod")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_ns("sidebar-mod")

filters
```

## Add and replace columns {.tabset}

Compute columns on your data using `shiynfilters::with_filter()` or 
`dplyr::mutate()`. Any new or replaced column receives a default filter input.

### shinyfilters {#columns-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_filter(dep_delay = -dep_delay, on_time = !delayed)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_filter(dep_delay = -dep_delay, on_time = !delayed)

filters
```

### dplyr {#columns-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(dep_delay = -dep_delay, on_time = !delayed)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(dep_delay = -dep_delay, on_time = !delayed)
```

## Use in a shiny app

You can place the filters directly inside a UI.

```{r, eval = FALSE}
library(bslib)
library(DT)
library(shiny)

ui <- page_sidebar(
	sidebar = sidebar(
		filters
	),
	DTOutput("data")
)
```

In the server, `shinyfilters_server()` to enable interdependency:
```{r, eval = FALSE}
server <- function(input, output, session) {
	sidebar <- shinyfilters_server(filters)
	output$data <- renderDT(datatable(sidebar$filtered))
}
```

Then run the app:
```{r, eval=FALSE}
shinyApp(ui, server)
```

## Notes

### Bookmarking

Apps that use bookmarking need `filterInput(filters)` inside their UI function
instead, so the inputs restore their bookmarked values. Placing `filters` in
their UI directly is an error.
