---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
```

```{r setup}
library(shinyfilters)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`shinyfilters()` turns a data frame into a set of filters. Printing it shows the
input each column gets, based on its type:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters()
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters()
```

## Set defaults for every column

Arguments to `shinyfilters()` apply to every column. `range = TRUE` uses a date
range for date columns, and `selectize = TRUE` uses selectize inputs in place
of select inputs:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE)
```

## Add a namespace

`with_ns()` prefixes every input id with a namespace, so the filters can be
used in a shiny module:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE) |>
	with_ns("sidebar-mod")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE) %>%
	with_ns("sidebar-mod")

filters
```

## Add and replace columns {.tabset}

Compute columns with dplyr's `mutate()`, or with `with_filter()`, which needs
no other package. A new column gets a filter based on its type; a replaced
column keeps its input:

### dplyr {#columns-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = requireNamespace("dplyr", quietly = TRUE), include = requireNamespace("dplyr", quietly = TRUE), message = FALSE}
library(dplyr)
```

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(dep_delay = -dep_delay, on_time = !delayed)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(dep_delay = -dep_delay, on_time = !delayed)
```

### shinyfilters {#columns-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filter(dep_delay = -dep_delay, on_time = !delayed)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filter(dep_delay = -dep_delay, on_time = !delayed)
```

## Choose inputs by keyword {.tabset}

The same functions set the input for individual columns. The shortest way is a
keyword like `"slider"`; see `?with_filter` for the full list. Name the columns
(`distance = "slider"`), or select them with tidyselect:

### dplyr {#inputs-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(across(where(is.numeric), "slider"))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(across(where(is.numeric), "slider"))
```

### shinyfilters {#inputs-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filter(where(is.numeric), "slider")
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filter(where(is.numeric), "slider")
```

## Choose inputs by shiny function {.tabset}

A shiny input function works in place of its keyword: `shiny::radioButtons`
gives the same input as `"radio"`:

### dplyr {#shiny-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(origin = shiny::radioButtons)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(origin = shiny::radioButtons)
```

### shinyfilters {#shiny-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filter(origin = shiny::radioButtons)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filter(origin = shiny::radioButtons)
```

## Choose inputs by custom function {.tabset}

Any other input function works too, such as `shinyWidgets::pickerInput`. It
receives the arguments shinyfilters computes for the column's type, here the
`choices` of a character column, so it must accept them:

### dplyr {#custom-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE) || !requireNamespace("shinyWidgets", quietly = TRUE)) "Install dplyr and shinyWidgets to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE)}
filters |>
	mutate(across(c(carrier, dest), shinyWidgets::pickerInput))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE)}
filters %>%
	mutate(across(c(carrier, dest), shinyWidgets::pickerInput))
```

### shinyfilters {#custom-shinyfilters}

`r if (!requireNamespace("shinyWidgets", quietly = TRUE)) "Install shinyWidgets to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE)}
filters |>
	with_filter(c(carrier, dest), shinyWidgets::pickerInput)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE)}
filters %>%
	with_filter(c(carrier, dest), shinyWidgets::pickerInput)
```

## Set an input's arguments {.tabset}

`as_filter()` pairs an input with arguments for it. They replace the ones
shinyfilters computes for the column, and an argument can use `.x`, the column
itself. Here a slider's `value` covers the column's range, which makes it a
range slider. `dep_delay` has missing values, so `range()` needs
`na.rm = TRUE`:

### dplyr {#args-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)
```

### shinyfilters {#args-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filter(
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filter(
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)
```

## Put it together {.tabset}

The steps chain into one pipeline, each seeing the columns the earlier ones
computed. `multiple = TRUE` is one more default: a select input that takes
several values starts with none, so the app below opens without filtering on
it:

### dplyr {#full-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights |>
	shinyfilters(
		range = TRUE,
		selectize = TRUE,
		multiple = TRUE
	) |>
	with_ns("sidebar-mod") |>
	mutate(
		on_time = !delayed,
		across(where(is.numeric), "slider"),
		origin = shiny::radioButtons,
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights %>%
	shinyfilters(
		range = TRUE,
		selectize = TRUE,
		multiple = TRUE
	) %>%
	with_ns("sidebar-mod") %>%
	mutate(
		on_time = !delayed,
		across(where(is.numeric), "slider"),
		origin = shiny::radioButtons,
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)

filters
```

### shinyfilters {#full-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters(
		range = TRUE,
		selectize = TRUE,
		multiple = TRUE
	) |>
	with_ns("sidebar-mod") |>
	with_filter(
		on_time = !delayed,
		across_filters(where(is.numeric), "slider"),
		origin = shiny::radioButtons,
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters(
		range = TRUE,
		selectize = TRUE,
		multiple = TRUE
	) %>%
	with_ns("sidebar-mod") %>%
	with_filter(
		on_time = !delayed,
		across_filters(where(is.numeric), "slider"),
		origin = shiny::radioButtons,
		dep_delay = as_filter(shiny::sliderInput, value = range(.x, na.rm = TRUE))
	)

filters
```

## Use in a shiny app

Place the filters in a UI to create their inputs.

```{r, eval = FALSE}
library(bslib)
library(DT)
library(shiny)

filters <- filters |> with_ns(NULL)

ui <- page_sidebar(
	sidebar = sidebar(
		filters
	),
	DTOutput("data")
)
```
<br>

Use `shinyfilters_server()` in the server to enable interdependency:
```{r, eval = FALSE}
server <- function(input, output, session) {
	sidebar <- shinyfilters_server(filters)
	output$data <- renderDT(datatable(sidebar$filtered))
}

shinyApp(ui, server)
```

`shinyfilters_server()` updates each input with the `update*Input()` function
that matches it. It knows the correct one for every shiny input `filterInput()`
creates. For any other input, such as `shinyWidgets::pickerInput`, name it with
`as_filter()`:

```{r, eval = FALSE}
as_filter(
	shinyWidgets::pickerInput,
	.update_fn = shinyWidgets::updatePickerInput
)
```

Apps that use bookmarking need `filterInput(filters)` inside their UI function
instead, so the inputs restore their bookmarked values. Placing `filters` in
their UI directly is an error.
