---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
library(shinyfilters)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`shinyfilters()` turns a data frame into a set of filters.

```{r}
filters <- shinyfilters(nyc_flights)
```

Printing it shows the input each column gets:
```{r}
filters
```

## Customizing filters

You can customize these settings on a `shinyfilters` object:

* A filter's input function
* The values used to create and update the input function
* Add a namespace

### Change a filter's input function

* Set an argument of `shinyfilters()`:
```{r}
shinyfilters(nyc_flights, range = TRUE)
```

* Use `with_defaults()`:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |> with_defaults(range = TRUE)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>% with_defaults(range = TRUE)
```

* Use `with_filters()`:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |> with_filters(date = "range")
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>% with_filters(date = "range")
```

* Use `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1", message = FALSE}
library(dplyr)

filters |> mutate(date = "range")
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1", message = FALSE}
library(dplyr)

filters %>% mutate(date = "range")
```
<br>

See `vignette("shinyfilters-catalog")` for the full list of input functions
available in `shinyfilters`.

### Change a filter's arguments

* Use `with_args()`:
```{r, eval = FALSE, include = getRversion() >= "4.1"}
filters |>
	with_args(dep_delay ~ value := range(.x))
```

```{r, eval = FALSE, include = !getRversion() >= "4.1"}
filters %>%
	with_args(dep_delay ~ value := range(.x))
```

* Use `as_filter()` without an input function argument:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filters(dep_delay = as_filter(value = range(.x)))
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filters(dep_delay = as_filter(value = range(.x)))
```

* Use `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	mutate(dep_delay = as_filter(value = range(.x)))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	mutate(dep_delay = as_filter(value = range(.x)))
```
<br>

### Change multiple filters together

* Set multiple filters inside `with_filters()`:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filters(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filters(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```

* Use `across_filters()` inside `with_filters()`:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filters(
		across_filters(where(is.numeric), as_filter(value = range(.x)))
	)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filters(
		across_filters(where(is.numeric), as_filter(value = range(.x)))
	)
```

* Set multiple filters inside `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	mutate(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	mutate(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```

* Use `dplyr::across()` inside `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	mutate(
		across(where(is.numeric), as_filter(value = range(.x)))
	)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	mutate(
		across(where(is.numeric), as_filter(value = range(.x)))
	)
```

* Use `dplyr::transmute()` to keep only the filters it names:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	transmute(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	transmute(
		dep_delay = as_filter(value = range(.x)),
		distance  = 1.609344 * distance
	)
```
<br>

### Use a namespace

* Set the `ns` argument of `shinyfilters()`:
```{r}
shinyfilters(nyc_flights, ns = "sidebar-mod")
```

* Use `with_ns()`:
```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_ns("sidebar-mod")
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_ns("sidebar-mod")
```

* Use `with_ns()` inside `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	mutate(with_ns("sidebar-mod"))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	mutate(with_ns("sidebar-mod"))
```
<br>

### Use a non-shiny input function

* Use `as_filter()` inside `with_filters()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1", message = FALSE}
library(shinyWidgets)

filters |>
	with_filters(dest = as_filter(pickerInput, .update_fn = updatePickerInput))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1", message = FALSE}
library(shinyWidgets)

filters %>%
	with_filters(dest = as_filter(pickerInput, .update_fn = updatePickerInput))
```

* Use `as_filter()` inside `dplyr::mutate()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = getRversion() >= "4.1"}
filters |>
	mutate(dest = as_filter(pickerInput, .update_fn = updatePickerInput))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE) && requireNamespace("shinyWidgets", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>%
	mutate(dest = as_filter(pickerInput, .update_fn = updatePickerInput))
```
<br>

## Selecting filters

After creating your filters, you can select multiple filters or a single filter,
allowing you to place the filters anywhere in an application

### Select multiple filters

* Use `[`:
```{r}
filters[starts_with("d")]
```

* Use `dplyr::select()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |> select(starts_with("d"))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>% select(starts_with("d"))
```
<br>

### Select one filter
<br>

* Use `$`:
```{r}
filters$dest
```
<br>

* Use `[[`:
```{r}
filters[["dest"]]
```
<br>

* Use `dplyr::pull()`:
```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1"}
filters |> pull(dest)
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1"}
filters %>% pull(dest)
```
<br>

## Use in a shiny app

### UI

You can place the filters directly inside a UI.

```{r, eval = FALSE}
library(bslib)
library(DT)
library(shiny)

ui <- page_sidebar(
	sidebar = sidebar(
		filters
	),
	DTOutput("data")
)
```

#### Organizing

To organize filters, you can select them individually or by group. Use `$` or
`[[` inside `accordion()`: it inspects its contents, so it needs the inputs
themselves, which `[` doesn't create.

```{r, eval = FALSE}
library(bslib)
library(DT)
library(shiny)

ui <- page_sidebar(
	sidebar = sidebar(
		accordion(
			layout_columns(
				class = "pe-3 pb-2 ps-3",
				filters$date
			),			
			accordion_panel(
				"Airports",
				filters[[c(origin, dest)]]
			),
			accordion_panel(
				"Time",
				filters[[contains("delay")]]
			),
			accordion_panel(
				"Other",
				filters[[!c(date, origin, dest, contains("delay"))]]
			)
		)
	),
	DTOutput("data")
)
```

### Server

In the server, use `shinyfilters_server()` to enable interdependency:
```{r, eval = FALSE}
server <- function(input, output, session) {
	sidebar <- shinyfilters_server(filters)
	output$data <- renderDT(datatable(sidebar$filtered))
}
```

### Run the app

Run the application as you normally would.
```{r, eval=FALSE}
shinyApp(ui, server)
```

## Notes

### Bookmarking

Apps that use bookmarking need `filterInput(filters)` inside their UI function
instead, so the inputs restore their bookmarked values. Placing `filters` in
their UI directly is an error.

### Custom objects

If you're filtering a custom object, you may need to extend shinyfilters. See
`vignette("extending-shinyfilters")` for details.
