---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
```

```{r setup}
library(shinyfilters)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`shinyfilters()` turns a data frame into a set of filters. Printing it shows the
input each column gets, based on its type:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters()
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters()
```

## Set defaults for every column

Arguments to `shinyfilters()` apply to every column. `range = TRUE` uses a date
range for date columns, and `selectize = TRUE` uses selectize inputs in place
of select inputs:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE)
```

## Add a namespace

`with_ns()` prefixes every input id with a namespace, so the filters can be
used in a shiny module:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE) |>
	with_ns("sidebar-mod")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE) %>%
	with_ns("sidebar-mod")

filters
```

```{r, include = FALSE}
filters_ns <- filters
```

## Add and replace columns {.tabset}

Compute columns with dplyr's `mutate()`, or with `with_filter()`, which needs
no other package. A new column gets a filter based on its type; a replaced
column keeps its input:

### dplyr {#columns-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = requireNamespace("dplyr", quietly = TRUE), include = requireNamespace("dplyr", quietly = TRUE), message = FALSE}
library(dplyr)
```

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- filters |>
	mutate(dep_delay = -dep_delay, on_time = !delayed)

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- filters %>%
	mutate(dep_delay = -dep_delay, on_time = !delayed)

filters
```

```{r, include = FALSE}
filters_dplyr <- filters
filters <- filters_ns
```

### shinyfilters {#columns-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- filters |>
	with_filter(dep_delay = -dep_delay, on_time = !delayed)

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- filters %>%
	with_filter(dep_delay = -dep_delay, on_time = !delayed)

filters
```

## Choose inputs per column {.tabset}

The same functions set the input for individual columns. Inputs are keywords
like `"slider"` or shiny functions like `shiny::sliderInput`; see
`?with_filter` for the full list. Name the columns (`distance = "slider"`), or
select them with tidyselect:

### dplyr {#inputs-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, include = FALSE}
filters_shinyfilters <- filters
filters <- filters_dplyr
```

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters |>
	mutate(across(where(is.numeric), "slider"))
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters %>%
	mutate(across(where(is.numeric), "slider"))
```

### shinyfilters {#inputs-shinyfilters}

```{r, include = FALSE}
filters <- filters_shinyfilters
```

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters |>
	with_filter(where(is.numeric), "slider")
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters %>%
	with_filter(where(is.numeric), "slider")
```

## Put it together {.tabset}

The steps chain into one pipeline:

### dplyr {#full-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE) |>
	with_ns("sidebar-mod") |>
	mutate(dep_delay = -dep_delay, on_time = !delayed) |>
	mutate(across(where(is.numeric), "slider"))

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE) %>%
	with_ns("sidebar-mod") %>%
	mutate(dep_delay = -dep_delay, on_time = !delayed) %>%
	mutate(across(where(is.numeric), "slider"))

filters
```

### shinyfilters {#full-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters(range = TRUE, selectize = TRUE) |>
	with_ns("sidebar-mod") |>
	with_filter(dep_delay = -dep_delay, on_time = !delayed) |>
	with_filter(where(is.numeric), "slider")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters(range = TRUE, selectize = TRUE) %>%
	with_ns("sidebar-mod") %>%
	with_filter(dep_delay = -dep_delay, on_time = !delayed) %>%
	with_filter(where(is.numeric), "slider")

filters
```

## Use in a shiny app

Place the filters in a UI to create their inputs. With the namespace, their
ids are prefixed (`sidebar-mod-origin`), so a module with the id
`"sidebar-mod"` reads them as `input$origin`:

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
	sidebarLayout(
		sidebarPanel(filters),
		mainPanel()
	)
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

Apps that use bookmarking need `filterInput(filters)` inside their UI function
instead, so the inputs restore their bookmarked values. Placing `filters` in
their UI directly is an error.
