---
title: "Get started with shinyfilters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with shinyfilters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>"
)
```

```{r setup}
library(shinyfilters)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
library(magrittr)
```

shinyfilters creates shiny inputs for filtering data. This article uses
`nyc_flights`, a sample of flights that left New York City in 2013.

## Filters from a data frame

`shinyfilters()` turns a data frame into a set of filters. Printing it shows the
input each column gets, based on its type:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters()
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters()
```

Arguments to `shinyfilters()` apply to every column. For example, `slider = TRUE`
uses sliders for numeric columns:

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
nyc_flights |>
	shinyfilters(slider = TRUE)
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
nyc_flights %>%
	shinyfilters(slider = TRUE)
```

## Choose inputs per column

Set the input for individual columns with dplyr's `mutate()`, or with
`with_filter()`, which needs no other package. Inputs are keywords like
`"radio"` or shiny functions like `shiny::radioButtons`. See `?with_filter` for
the full list.

### Name or select columns {.tabset}

Name the columns, or select them with tidyselect:

#### dplyr {#columns-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = requireNamespace("dplyr", quietly = TRUE), include = requireNamespace("dplyr", quietly = TRUE), message = FALSE}
library(dplyr)
```

```{r, eval = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights |>
	shinyfilters() |>
	mutate(origin = "radio", carrier = "selectize") |>
	mutate(across(where(is.numeric), "slider"))

filters
```

```{r, eval = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE), include = !getRversion() >= "4.1" && requireNamespace("dplyr", quietly = TRUE)}
filters <- nyc_flights %>%
	shinyfilters() %>%
	mutate(origin = "radio", carrier = "selectize") %>%
	mutate(across(where(is.numeric), "slider"))

filters
```

#### shinyfilters {#columns-shinyfilters}

```{r, eval = getRversion() >= "4.1", include = getRversion() >= "4.1"}
filters <- nyc_flights |>
	shinyfilters() |>
	with_filter(origin = "radio", carrier = "selectize") |>
	with_filter(where(is.numeric), "slider")

filters
```

```{r, eval = !getRversion() >= "4.1", include = !getRversion() >= "4.1"}
filters <- nyc_flights %>%
	shinyfilters() %>%
	with_filter(origin = "radio", carrier = "selectize") %>%
	with_filter(where(is.numeric), "slider")

filters
```

### Select and name columns in one call {.tabset}

`across()` and `across_filters()` select columns and name one input for all of
them, alongside named columns:

#### dplyr {#across-dplyr}

`r if (!requireNamespace("dplyr", quietly = TRUE)) "Install dplyr to see this example."`

```{r, eval = requireNamespace("dplyr", quietly = TRUE), include = requireNamespace("dplyr", quietly = TRUE)}
mutate(
	filters,
	across(where(is.character), "selectize"),
	origin = "radio"
)
```

#### shinyfilters {#across-shinyfilters}

```{r}
with_filter(
	filters,
	across_filters(where(is.character), "selectize"),
	origin = "radio"
)
```

## Use in a shiny app

Place the filters in a UI to create their inputs:

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
	sidebarLayout(
		sidebarPanel(filters),
		mainPanel()
	)
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

Apps that use bookmarking need `filterInput(filters)` inside their UI function
instead, so the inputs restore their bookmarked values.
